Optifcal Observations of an Extraordinarily Young Type Ia Supernova iPTF16abc

### 2017-02-01 doc/ms.tex
Full changes to the language in the entire manuscript.

More commonly, the only observed light curve of a SN Ia is purely
powered by the radioactive decay of synthesized $^{56}$Ni. Since
$^{56}$Ni atoms are synthesized, mixed and deposited into different
layers inside the ejecta, the SN may experience a dark period after
the SN shock breakout and before the radioactive energy diffuses to
the photosphere \citep{2014ApJ...784...85P}. In the case of strong
mixing, $^{56}$Ni energy can reach the photosphere rapidly. As such,
the dark period is very short and the light curve shows a fast initial
rise. In the opposite case, it may take up to a couple of days for
the $^{56}$Ni energy to diffuse to the photosphere. The initial rise
of the light curve is then moderate \citep{2016ApJ...826...96P}. To
summarize, the initial rise of the light curve of a SN Ia conveys
information on distribution of synthesized $^{56}$Ni.

In this paper, we report observations of an extraordinarily young SN
Ia iPTF16abc, which was discovered by the intermediate Palomar
Transient Factory on 2016 April $3.36$\footnote{all times in this
  paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g$-band magnitude of
$21.31\pm0.27$ \citep{2016PASP..128k4502C,2016ATel.8907....1M}. The
transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221 at 100\,Mpc. No activity was detected at the same location
down to $g=22.1$\,mag on April $2.42$. Our spectroscopic follow-up
compaign classified iPTF16abc as a normal SN Ia
\citep{2016ATel.8909....1C}.

This paper is organized as follows: Section \ref{sec:obs} describes
photometric and spectroscopic observations of iPTF16abc. Section
\ref{sec:usual_staff} establishes that iPTF16abc is a normal SN Ia
in NGC\,5221. Section \ref{sec:first_light} analyzes the early
light curve and spectra.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{lightcurve.pdf}
  \caption{Multi-band light curves of iPTF16abc are shown. Filters are
    denoted by different colors and observation instruments by
    different markers. The $t_{max}$ time is the B-band maximum
    determined by SALT2 (Section \ref{sec:classification}). The solid
    curves are best-fit results from SALT2. The black ticks near the
    top of the figure shows epochs of spectroscopic observations.}
  \label{fig:lightcurve}
\end{figure*}

As part of the iPTF transient survey in the 2016 spring quarter, the
field of iPTF16abc was observed in $g$- or $R$-band every night by the
CFH12K camera \citep{2000SPIE.3965...58S} on the 48-inch telescope at
Palomar Observatory (P48). The images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN. The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

After discovery, we utilized the rainbow camera of the SED Machine
(\ycao{REF}) mounted on the 60-inch telescope at Palomar Observatory
(P60) to undertake photometric observations in $g$, $r$ and $i$
filters. The image differencing against the archival SDSS images and
forced PSF photometry on the subtracted images were performed by the
Fremling Automated Pipeline \citep{2016A&A...593A..68F}. The
photometry is also calibrated to the SDSS catalog \ycao{which DR?}.

Las Cambres Observatory Global Network (LCOGT) also carry out
photometric observations in the \textit{BVgri} filters with its 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{gri}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

In space, \textit{Swift} observed iPTF16abc for 14 epochs, covering
from the very early phase to the post-peak phase. Aperture photometry
are carried out on the images taken by its Ultraviolet-Optical
Telescope (UVOT) with the usual procedures in the HEASoft and
corrected for the coincident loss and aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. No pre-SN UVOT image at the SN location
is available in the \textit{Swift} archive.  Visual inspection to the
UVOT images suggests that the background galaxy light in the UVOT
filters is probably negligible. No X-ray emission was detected at the
location of the SN by the X-ray Telescope (XRT) in any of these
epochs.

The multi-color light curves of iPTF16abc are illustrated in Figure
\ref{fig:lightcurve}.  For convenience, magnitudes in \textit{all}
filters are in the AB system with a zero point of $3631$\,Jy.

Spectroscopic observations of iPTF16abc were undertaken with a variety
of telescopes and instruments in multiple epochs spanning from a
couple of days after explosion to two months after its maximum. An
observing log is listed in Table \ref{tab:spec_obs_log}. Data
reduction is undertaken with usual routines in IDL/Python. Except for
the high-resolution spectra obtained by VLT instruments, the optical
spectral evolution of iPTF16abc is illustrated in Figure
\ref{fig:spec_seq}.

@@

The foreground Galactic extinctioin along the directionof iPTF16abc
has $E(B-V)=0.0279\,\textrm{mag}$ \citep{2011ApJ...737..103S}.

In the highest-resolution spectrum of iPTF16abc by UVES, individual
components of both \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets show
double-absorption profiles (Figure \ref{fig:narrow_features}),
indicating two sources of absorption along the line of sight. Fitting
two Gaussian kernels to each line of the \ion{Na}{1} doublet
simultaneously leads to redshifts of $0.02313820\pm0.00000032$ and
$0.02322408\pm0.00000033$. The total equivalent widths of the
\ion{Na}{1}\,D1 and D2 lines are $0.595\pm0.009\,\textrm{\AA}$ and
$0.609\pm0.008\,\textrm{\AA}$m, respectively. Using the empirical
relation between the equivalent width of \ion{Na}{1}\,D lines and
reddening $E(B-V)$ \citep{2012MNRAS.426.1465P}, we derive
$E(B-V)=0.361\pm0.025\textrm{mag}$. We caution that the empirical
relation has a quite large scatter.

In the X-shooter spectra of iPTF16abc, we also identify narrow
absorption features \ion{K}{1}\,7665\,\AA and 7699\,\AA at consistent
redshifts. However, the resolution of the X-shooter spectra is not
fine enough to resolve the double-absorption profiles of \ion{Ca}{2}
and \ion{Na}{1}. The X-shooter spectra do not show the diffusive
interstellar band at 5780\,\AA as well.  Albeit some debates (e.g.,
\citealt{2013ApJ...779...38P}), the $E(B-V)$ value derived from
\ion{Na}{1}\,D absorption provides the best estimate in the case of
iPTF16abc.

@@

The \ion{Na}{1}\,D doublet are seen in multiple spectra spanning
from pre-peak to post-peak phases. Despite the instrumental widening of
different instrument configurations, we do not detect obvious variation
in the profiles of the doublet.

We also note in Figure \ref{fig:narrow_features} that the absorption
profiles of \ion{Ca}{2}\,H$+$K are quite different from those of \ion{Na}{1}\,D
doublet, implying that the dust compositions in the two absorption sources
may have some difference.

@@

We run Supernova Identification (SNID; \citealt{2007ApJ...666.1024B})
on the low-resolution spectrum of iPTF16abc at $+18.8$ and found best
matches to normal SNe Ia. In fact, the characteristic features of a SN
Ia, such as \ion{Si}{2}, \ion{S}{2}, can be easily identified in the
spectra of iPTF16abc (Figure \ref{fig:spec_seq}).

Next we standardize the light curve of iPTF16abc by feeding its P48
and P60 light curves into the \texttt{sncosmo} module\footnote{The
  \texttt{sncosmo} Python module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.4.x/}.} and fit a light
curve model, which composes of the SALT2 template
\citep{2007A&A...466...11G} modified by the line-of-sight extinction
curve \citep{1999PASP..111...63F} with $E(B-V)$ values from Section
\ref{sec:reddening} and $R_V=3.1$. We obtain the rest-frame
\textit{B}-band peak time $\textrm{MJD}_{max}=57499.65\pm0.02$, the coefficient
of the zeroth principle component $x_0=0.0275\pm0.0002$, the
coefficient of the first principle component $x_1=1.200\pm0.043$, and
the color term $c=-0.3353\pm0.0054$. The best-fit model also gives an
unreddened apparent peak magnitude of $m^*_{B}=14.4\,\textrm{mag}$ in
the SN rest frame.

@@

After establishing iPTF16abc as a normal SN ia, we use the latest
calibration \citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a
distance modulus $\mu=34.66\pm0.03\,\textrm{mag}$ to the SN, provided
that the host galaxy of iPTF16abc has a stellar mass less than
$10^{10}\sm$.

The location of iPTF16abc is spatially coincident with a tidal tail of
galaxy NGC\,5221. \citet{2007A&A...465...71T} derived a distance
modulus of $35.0\pm0.4\,\textrm{mag}$ from the Tully-Fisher
relation. This distance modulus is consistent with that of iPTF16abc.

Separately, \citet{1998A&AS..130..333T} observed the 21-cm line in
this galaxy and measured a redshift of $0.0233303\pm0.000027$.  The
two components in the \ion{Na}{1}\,D have a relative velocity of
$-57.6\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$ and
$-31.8\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$, suggesting that both
absorption resources are probaly located on the tidal tail of
NGC\,5221.

@@

In this section, we estimate the first light and explosion time of
iPTF16abc.

@@

Extrapolation of the earliest-phase light curve with a simple
mathematical model is usually used to estimate the first light time of
a SN. In theory, \citet{1982ApJ...253..785A} derived a quadratic law
for an ideal expanding fireball with a constant temperature. In order
to account for variation of photospheric temperatures during
expansion, we utilize a power-law curve to model the early light curve.
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ .
\end{equation}
Since the first few detections of the SN were made in the \textit{g}
band, our modeling is limited to the early \textit{g}-band light
curve. We minimize $\chi^2$ in a large search grid of $t_0$, $\alpha$
and the proportionality constant. We also note that the definition of
the time window for the early light curve may affect the fitting
result, so we experiment the fitting procedure with different time
windows. The modeling results show that the SN flux between
$t=-19\,\textrm{days}$ and $t=-15\,\textrm{days}$ rises approximately
linearly. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. With the
best-fit model of $\alpha=0.92$ and $t_0=-18.47$ days, our first
observation was made only $0.18\,\textrm{day}$ after the first light
of the SN. The total rise time to the \textit{B}-band peak is $18.47$
days.

@@

Starting from $t=-15\,\textrm{days}$, the \textit{g}-band light curve
rises significantly faster than the best-fit model above, indicating a
greater value of the power-law index $\alpha$. In fact, the light
curve between $t=-14$ and $t=-8$\,days can be fitted with another
power law of index $1.40$. In another word, the whole light curve
before $t=-8$\,days can be approximated by a broken power-law model
(e.g.,
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2016arXiv161202097Z,
  2016arXiv161202725Z}).

As discussed in Section \ref{sec:lc_energy}, the initial light curve
of iPTF16abc is probably purely powered by radioactive decay of
$^{56}$Ni. Therefore, the initial rise of the light curve depends on
the depth of the shallowest layer where $^{56}$Ni is deposited
\citep{2014ApJ...784...85P}. According to theoretical mdels
\citep{2016ApJ...826...96P}, the linearly rising light curve implies
strong mixing of $^{56}$Ni in the ejecta. As a result, the time lag
between the SN explosion and the initial rise of the radioactively
powered light curve is negligible.

@@

In order to constrain the explosion time of the SN independently of
the SN light curve, \citet{2014ApJ...784...85P} suggested to measure
the expansion velocities of the photosphere, because the expansion of
the ejecta starts at the onset of the SN explosion. Assuming a
constant opacity in the ejecta, the photospheric velocity evolves as
$v_{ph}\propto(t-t_{exp})^{-0.22}$. While the photospheric velocities
are not easy to measure, line velocities of \ion{Si}{2} or \ion{Ca}{2}
are usually used to approximate the photospheric velocities
\citep{2014ApJ...784...85P,2016ApJ...826..144S}.

In the case of iPTF16abc, the \ion{Ca}{2} IR triplet is very weak in
the spectra probably due to high temperatures in the ejecta. Thus we
perform our measurements on the \ion{Si}{2}\,6355 line. Visual
inspection of the spectra shows no sign of multi-velocity components
of \ion{Si}{2}, and that the \ion{C}{2}\,6580 line overlaps the red
wing of the \ion{Si}{2} line. Consequently we apply a mathematical
model of two gaussian kernels superposed on a linear term, which
accounts for the absorption of \ion{Si}{2} and \ion{C}{2} and the
continuum component, respectively, to approach the observed spectra
between $5,900$ and $6,500$\,\AA (rest-frame). The expansion
velocity of \ion{Si}{2} is calculated at the central wavelength
of the \ion{Si}{2} Gaussian kernel.

@@

We fit the measured velocities of \ion{Si}{2}\,6355 line to the
$v\propto(t-t_{exp})^{-0.22}$ model by minimizing the $\chi^2$ value
and find the best-fit explosion time $t_{exp}=-17.95\,\textrm{days}$
with a 3-$\sigma$ confidence interval between $-17.4\,\textrm{days}$
and $-18.3\,\textrm{days}$ (Figure \ref{fig:velocity_t_exp}). We also
alter the power-law index to $-0.20$ and $-0.24$ to examine how
sensitively the result depends on the assumed power-law index.  we
find consistent results within respective 3-$\sigma$ confidence
intervals.

Keeping in mind that $t_{exp}$ has a very large uncertainty mostly due
to the assumption in the $v\propto(t-t_{exp})^{-0.22}$ model, we
compare the estimated $t_{exp}$ to $t_0$ (left panel of Figure
\ref{fig:velocity_t_exp}) and find that $t_0\lessapprox t_{exp}$.
Since physical causality requires $t_{exp}<t_0$, we draw a qualitative
conclusion that $t_0\simeq t_{exp}$. This conclusion is consistent
with our inference from the early light curve analysis in Section
\ref{sec:lc_fit}.

With $t_{exp}$, we estimate the actual rise time of iPTF16abc from the
SN explosion to its \textit{B}-band maximum to be $17.95$ days.  In
comparison, SN2011fe has $t_0=-17.7$ days \citep{2013A&A...554A..27P}
with a preceding dark period of $\sim 1$ day
\citep{2014ApJ...784...85P}. ASASSN-14lp has $t_0=-16.94$ days with a
preceding dark period of about acouple of days
\citep{2016ApJ...826..144S}. This comparison suggests that the actual
rise time of a SN Ia between the SN explosion and the \textit{B}-band
maximum is roughly a constant, which largely depends on the total mass
of synthesized $^{56}$Ni.

@@

Another intriguing spectral feature of iPTF16abc is strong absorption of
\ion{C}{2}\,6580 and \ion{C}{2}\,7234 lines. Our analysis on the
\ion{Si}{2}\,6355 line also allows us to measure the velocities and
pseudo-equivalent widths (pEWs) of \ion{C}{2}\,6580. We also fit a Gaussian
kernel superposing on a linear term to the spectral region near
the \ion{C}{2}\,7234 line to measure its velocity and pseudo-equivalent
width. The velocity evolution of \ion{C}{2} is shown in the right panel
of Figure \ref{fig:velocity_t_exp} and the pEW evolution
in Figure \ref{fig:ew}.

Our measurements show that the strong \ion{C}{2} absorption
only appears in the very early phases. In the spectrum taken at
$t=-15.8$ days, the pEW of \ion{C}{2}\,6580 is comparable to that
of \ion{Si}{2}\,6355. Then the pEW of \ion{C}{2} decreases
until they disappear at about $t=-10$ days.

@@

Formation of \ion{C}{2} absorption requires existence of both unburned
carbon and $^{56}$Ni, the latter of which is responsible to ionize the
carbon atoms. The strong and short-lived \ion{C}{2} features observed
in the spectra of iPTF16abc show that a certain amount of carbon atoms
exist in the outer layers of the ejecta and that $^{56}$Ni atoms are also
strongly mixed in the same layers. This provides a third piece of evidence
to strong mixing in the iPTF16abc ejecta.

>>>>

The detection of SN shock breakout or an ejecta-companion collision is exceedingly rare, and the vast majority of SNe Ia are observed to be 
powered purely by the radioactive decay of $^{56}$Ni. 
SNe Ia may experience a dark period after
the SN shock breakout but before radioactive energy diffuses into
the photosphere depending on how the newly synthesized $^{56}$Ni is mixed and deposited into different layers of the ejecta \citep{2014ApJ...784...85P}. If there is strong
mixing, photons from $^{56}$Ni decay can reach the photosphere rapidly. This leads to a short dark period and a fast initial rise in the observed optical light curve. For weak mixing, it may take several days for
the radioactive energy to diffuse to the photosphere. The initial rise
of the light curve is then moderate \citep{2016ApJ...826...96P}. Thus, the initial rise of SN Ia light curves conveys
information on the distribution of synthesized $^{56}$Ni.

In this paper, we report observations of an extraordinarily young SN
Ia, iPTF16abc, which was discovered by the intermediate Palomar
Transient Factory on 2016 April $3.36$\footnote{All times in this
  paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g$-band magnitude of
$21.31\pm0.27$ \citep{2016PASP..128k4502C,2016ATel.8907....1M}. The
transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221, which lies at a distance of $\sim$100\,Mpc. iPTF16abc is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery. Our spectroscopic follow-up
compaign classified iPTF16abc as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis provide multiple lines of evidence that the $^{56}$Ni produced by iPTF16abc was strongly mixed throughout the ejecta. 
\ycao{My personal preference is to not include section by section outline, esp for short papers - this info is clear from section titles. But if you want to include this feel free}

% This paper is organized as follows: Section \ref{sec:obs} describes
% photometric and spectroscopic observations of iPTF16abc. Section
% \ref{sec:usual_staff} establishes that iPTF16abc is a normal SN Ia
% in NGC\,5221. Section \ref{sec:first_light} analyzes the early
% light curve and spectra.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.95\textwidth]{lightcurve.pdf}
  \caption{Multi-band light curves of iPTF16abc. Each filter is
  represented with a different color, and each instrument with a 
  different symbol. $t_{max}$ is the time of B-band maximum
    determined by SALT2 (Section \ref{sec:classification}). The solid
    curves are best-fit results from SALT2. The black ticks near the
    top of the figure show epochs of spectroscopic observations.}
  \label{fig:lightcurve}
\end{figure*}

During the spring of 2016, the iPTF survey observed the field of iPTF16abc every night in either the $g$- or $R$-band.
Survey observations were conducted with the
CFH12K camera \citep{2000SPIE.3965...58S} on the Palomar Observatory 48-inch telescope 
 (P48). Images were processed by the IPAC image
subtraction and discovery pipeline which subtracts off the background
galaxy light with stacked pre-SN images and performs forced
point-spread-function (PSF) photometry at the location of the SN.\ycao{Add a reference to Frank's PTFIDE paper - http://iopscience.iop.org/article/10.1088/1538-3873/129/971/014002/pdf} The
photometry is then calibrated to the PTF photometric catalog
\citep{2012PASP..124..854O}.

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDM; \ycao{REF}) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline \citep{2016A&A...593A..68F} to subtract galaxy light from the SEDM images using achival Sloan Digital Sky Survey (SDSS) images as reference images. This pipeline then performed forced-PSF photometry at the location of iPTF16abc, which is calibrated to the SDSS catalog \ycao{which DR?}.

photometric observations in the \textit{BVgri} were conducted by the Las Cambres Observatory Global Network (LCOGT) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{gri}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

In space, \textit{Swift} observed iPTF16abc on 14 epochs, beginning $\sim$15 d pre-maximum light through $\sim$5 d post maximum. \ycao{check and confirm numbers} The SN flux is measured via aperture photometry on Ultraviolet-Optical
Telescope (UVOT) images via the usual procedures in the HEASoft, including corrections for the coincident loss and aperture loss. The image counts
are converted to physical fluxes using the latest calibration
\citep{2011AIPC.1358..373B}. There are no pre-SN UVOT images at the SN location available in the \textit{Swift} archive.  Visual inspection of the
UVOT images suggests negligible host-galaxy contamination in our UVOT flux measurements. There is no X-ray emission detected at the 
SN position by the \textit{Swift} X-ray Telescope (XRT).

The multi-color light curves of iPTF16abc are illustrated in 
Figure~\ref{fig:lightcurve}.  For convenience, magnitudes in \textit{all}
filters are in the AB system with a zero point of $3631$\,Jy.

Spectroscopic observations of iPTF16abc were undertaken with a variety
of telescopes and instruments in multiple epochs spanning from a
couple of days after explosion to two months after $B$-band maximum. An
observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}. The optical
spectral evolution of iPTF16abc is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

The foreground Galactic extinction toward iPTF16abc
is $E(B-V)=0.0279\,\mathrm{mag}$ \citep{2011ApJ...737..103S}.

Our highest-resolution spectrum of iPTF16abc, obtained with VLT/UVES, shows multiple absorption components for each of the individual lines in the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets (Figure \ref{fig:narrow_features}). This indicates there are two sources of absorption along the line of sight. Fitting
two Gaussian kernels to each line of the \ion{Na}{1} doublet
simultaneously leads to redshifts of $0.02313820\pm0.00000032$ and
$0.02322408\pm0.00000033$. The total \ion{Na}{1}\,D equivalent width (EW) is $0.595\pm0.009\,\textrm{\AA}$ and
$0.609\pm0.008\,\textrm{\AA}$, respectively. Using the empirical
relation between the EW of \ion{Na}{1}\,D lines and
reddening $E(B-V)$ \citep{2012MNRAS.426.1465P}, we derive
$E(B-V)=0.361\pm0.025\textrm{mag}$. We note that this relation has large scatter.

In the VLT/X-shooter spectra of iPTF16abc, we also identify narrow \ion{K}{1}\,7665\,\AA\ and 7699\,\AA\ absorption at the same
redshifts. The double-absorption profiles of \ion{Ca}{2}
and \ion{Na}{1} are not resolved in the VLT/X-shooter spectra. Furthermore, the X-shooter spectra do not show the diffusive
interstellar band at 5780\,\AA. For iPTF16abc, \ion{Na}{1}\,D provides the best estimate of host-galaxy reddening (though see \citealt{2013ApJ...779...38P} for a discussion on why \ion{Na}{1}\,D is a poor tracer of redenning).

@@

We detect \ion{Na}{1}\,D in multiple spectra both before and after peak. After accounting for differing instrumental resolutions, we do not detect any obvious variations in the \ion{Na}{1}\,D profiles.

Finally, we note that the \ion{Ca}{2}\,H$+$K absorption profile is quite different from the \ion{Na}{1}\,D profile. 
This suggests that the dust composition may differ in the two absorption components.

@@

Using the Supernova Identification (SNID; \citealt{2007ApJ...666.1024B}) package, 
we find the low-resolution spectrum of iPTF16abc at $+18.8$ d is best matched by normal SNe Ia. Several characteristic features of a SN
Ia, such as \ion{Si}{2}, \ion{S}{2}, can be easily identified in the
spectra of iPTF16abc (Figure \ref{fig:spec_seq}).

To determine the brightness and time of $B$-band maximum for 
iPTF16abc, we fit the P48 and P60 light curves with the \texttt{sncosmo} software package.\footnote{The
  \texttt{sncosmo} Python module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.4.x/}.} This fit includes a SALT2 template \citep{2007A&A...466...11G} modified by the line-of-sight extinction
curve \citep{1999PASP..111...63F} with $E(B-V)$ values from Section
\ref{sec:reddening} and $R_V=3.1$.

We determine the time of rest-frame \textit{B}-band maximum to be 
 $\textrm{MJD}_{max}=57499.65\pm0.02$, the coefficient
of the zeroth principle component $x_0=0.0275\pm0.0002$, the
coefficient of the first principle component $x_1=1.200\pm0.043$, and
the color term $c=-0.3353\pm0.0054$. The best-fit model also gives an
unreddened apparent peak magnitude of $m^*_{B}=14.4\,\textrm{mag}$ \ycao{uncertainties on Bmax?} in
the SN rest frame.

@@

After establishing iPTF16abc as a normal SN Ia, we use the latest
calibration \citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a
distance modulus $\mu=34.66\pm0.03\,\textrm{mag}$ to the SN, provided
that the host galaxy of iPTF16abc has a stellar mass less than
$10^{10}\sm$.

The location of iPTF16abc is spatially coincident with a tidal tail of
galaxy NGC\,5221. \citet{2007A&A...465...71T} derived a distance
modulus of $35.0\pm0.4\,\textrm{mag}$ to NGC\,5221 from the Tully-Fisher
relation. This distance modulus is consistent with that of iPTF16abc.

Separately, \citet{1998A&AS..130..333T} observed the 21-cm line in
this galaxy and measured a redshift of $0.0233303\pm0.000027$.  The
two components in the \ion{Na}{1}\,D have a relative velocity of
$-57.6\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$ and
$-31.8\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$, suggesting that both
absorption resources are probaly located on the tidal tail of
NGC\,5221.\ycao{This is confusing - are those velocities relative to host redshift or LSR?}

@@

@@

The time of first light for SNe is usually estimated by 
extrapolating early-phase light curves to match simple 
mathematical models. Assuming an ideal exapnding fireball with 
constant temperature, \citet{1982ApJ...253..785A} derives that $f 
\propto t^2$, where $f$ is the SN flux and $t$ is the time since 
explosion. To account for variations of the photospheric 
temperature during expansion, we model the early light curve as a power law:
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
where $t_0$ is the time of first light and $alpha$ is the power-law index. We limit our fit to the earliest detections of iPTF16abc, which were exclusively in $g$-band. To determine the best fit parameters, we search a large grid over $t_0$, $\alpha$
and the proportionality constant, and minimize $\chi^2$. 
We note that the definition of
the time window for the early light curve may affect the fitting
result, so we experiment the fitting procedure with different time
windows. \ycao{I don't understand what this previous sentence means}
The modeling results show that the SN flux rises approximately
linearly between
$t=-19\,\textrm{d}$ and $t=-15\,\textrm{d}$. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. With the
best-fit model of $\alpha=0.92$ and $t_0=-18.47$ days, our first
detection of iPTF16abc was made only $\sim{0.2}\,\textrm{d}$ after 
the SN first light. From the fit, we additionally infer a total rise time to \textit{B}-band peak of $18.47$
days.

@@

Starting around $t=-15\,\textrm{d}$, the $g$-band light curve 
deviates from the above best-fit model and 
rises significantly faster, indicating a
change in the power-law index $\alpha$. In fact, between $t=-14\,\mathrm{d}$ and $t=-8\,\mathrm{d}$ the light curve is well fit by a power law of index $1.40$. In other words, the entire light curve
before $t=-8$\,days can be approximated by a broken power-law model
(e.g.,
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2016arXiv161202097Z,
  2016arXiv161202725Z}).

As discussed in Section \ref{sec:lc_energy}, the initial light curve
of iPTF16abc is probably purely powered by the radioactive decay of
$^{56}$Ni. Thus, the initial rise of the light curve depends on
the depth of the shallowest layer where $^{56}$Ni is deposited
\citep{2014ApJ...784...85P}. According to theoretical mdels
\citep{2016ApJ...826...96P}, the linearly rising light curve implies
strong mixing of $^{56}$Ni in the ejecta. As a result, the time lag
between the time of explosion and SN first light is negligible.

@@

Due to the possibility of an early dark phase for SNe Ia, \citet{2014ApJ...784...85P} suggest that photospheric velocity measurements be used to determine the time of explosion, because the ejecta expand from the onset of the SN explosion. Assuming a
constant opacity in the ejecta, \citeauthor{2014ApJ...784...85P} find that the photospheric velocity evolves as
$v_{ph}\propto(t-t_{exp})^{-0.22}$. While the photospheric velocity
is not easy to measure, line velocities of \ion{Si}{2} or \ion{Ca}{2}
can be used as a proxy
\citep{2014ApJ...784...85P,2016ApJ...826..144S}.

In the case of iPTF16abc, the \ion{Ca}{2} IR triplet is very weak, 
probably due to high ejecta temperature. Thus, we
determine the photospheric velocity from the \ion{Si}{2}\,6355 line. Visual
inspection shows no sign of multi-velocity components
of \ion{Si}{2}, and that the \ion{C}{2}\,6580 line overlaps the red
wing of the \ion{Si}{2} line. Consequently, we model the observed 
spectra between $5,900$ and $6,500$\,\AA (rest-frame)
with two gaussian kernels superposed on a linear baseline, which
accounts for \ion{Si}{2}, \ion{C}{2} and the
continuum, respectively. The expansion
velocity of \ion{Si}{2} is measured by the central wavelength
of the \ion{Si}{2} Gaussian kernel.

@@

We fit the measured velocities of \ion{Si}{2}\,6355 line to the
$v\propto(t-t_{exp})^{-0.22}$ model by minimizing the $\chi^2$ value
and find the best-fit explosion time $t_{exp}=-17.95\,\textrm{days}$
with a 3-$\sigma$ confidence interval between $-17.4\,\textrm{days}$
and $-18.3\,\textrm{days}$ (Figure \ref{fig:velocity_t_exp}). We additionally
alter the power-law index to $-0.20$ and $-0.24$ to examine the
sensitively of the result on the assumed power-law index.  This 
leads to consistent results within the respective 3-$\sigma$ 
confidence intervals.

Compare the estimated $t_{exp}$ to $t_0$ (left panel of Figure
\ref{fig:velocity_t_exp}), we find that $t_0\lessapprox t_{exp}$.\footnote{Note that the uncertainties on $t_{exp}$ are large due to assumptions in the $v\propto(t-t_{exp})^{-0.22}$ model.}
Since physical causality requires $t_{exp}<t_0$, we draw the qualitative
conclusion that $t_0\simeq t_{exp}$, which is consistent
with our inference from the early light curve analysis (Section
\ref{sec:lc_fit}).

From $t_{exp}$, we estimate the actual rise time of iPTF16abc from the
time of explosion to \textit{B}-band maximum to be $17.95$ d.  In
comparison, SN2011fe has $t_0=-17.7$ d \citep{2013A&A...554A..27P}
with a preceding dark period of $\sim 1$ day
\citep{2014ApJ...784...85P}. ASASSN-14lp has $t_0=-16.94$ days with a
preceding dark period of about acouple of days
\citep{2016ApJ...826..144S}. This comparison suggests that the actual
rise time of a SN Ia between the SN explosion and the \textit{B}-band
maximum is roughly a constant, which largely depends on the total mass
of synthesized $^{56}$Ni.\ycao{I am not convinced by this last sentence}

@@

iPTF16abc exhibits unusually strong absorption of
\ion{C}{2} $\lambda\lambda$6580, 7234 for SNe Ia. Our analysis of the
\ion{Si}{2}\,6355 line also allows us to measure the velocities and
pseudo-equivalent widths (pEWs) of \ion{C}{2}\,6580. We also fit a Gaussian
kernel with linear baseline to the spectral region near
the \ion{C}{2}\,7234 line to measure its velocity and pEW. 
The velocity evolution of \ion{C}{2} is shown in the right panel
of Figure \ref{fig:velocity_t_exp} and the pEW evolution
in Figure \ref{fig:ew}.

Our measurements show that the strong \ion{C}{2} absorption
only appears in the very early phases. In the spectrum taken at
$t=-15.8$ days, the pEW of \ion{C}{2}\,6580 is comparable to that
of \ion{Si}{2}\,6355. Subsequent spectra the pEW of \ion{C}{2} 
decreases until the feature is no longer detectable 
around $t=-10$ days.

@@

\ion{C}{2} absorption requires the existence of both unburned
carbon and $^{56}$Ni, which radioactively decays and ionizes the C 
atoms. The strong and short-lived \ion{C}{2} features reveal 
that $^{56}$Ni atoms have been strongly mixed in to the outter 
layers of the ejecta. This provides a third piece of evidence
for strong mixing in the iPTF16abc ejecta.

Carbon signatures are seen in $> 1/4$ of all normal SNe Ia before
maxima
\citep{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T},
but the signatures are usually weak. Even in SN2011fe, which was discovered and spectroscopically observed shortly after explosion, the \ion{C}{2}
features are not strong in the first spectra
\citep{2012ApJ...752L..26P}.  The only normal SN Ia known to have
strong \ion{C}{2} features at early phases is SN2013dy
\citep{2013ApJ...778L..15Z}. Unlike iPTF16abc, however, the pEWs 
of the \ion{C}{2} features in SN2013dy are \ycao{not?} as large as 
those of \ion{Si}{2}\,6355 in the same spectra.

\subsection{Discussion}
\label{sec:lc_energy}

Multiple energy sources can contribute to the early radiation from a SN Ia: SN shock
breakout, SN ejecta-companion collision, and radioactive 
$^{56}$Ni. The previous analysis is based on the assumption that the light curve of
iPTF16abc is powered purely by radioactive decay. Here we discuss the
other two possibilities.
